---
title: "PS3"
output: pdf_document
author:Saniya Ambavanekar
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown



```{r}
library(NHANES)
library(ggplot2)
men_data=subset(NHANES,NHANES$Gender=="male")
women_data=subset(NHANES,NHANES$Gender=="female")

```

Observing the trend of systolic blood pressure with the age of men
```{r}
Age_men_data=men_data[c("Age","BPSysAve")]
Age_men_data=na.omit(Age_men_data)

Age_women_data=women_data[c("Age","BPSysAve")]
Age_women_data=na.omit(Age_women_data)
```

Observing the trend by fitting curve using the loess method
```{r}
Age_men.gg=ggplot(Age_men_data,aes(x=Age_men_data$Age,y=Age_men_data$BPSysAve))+geom_point()
Age_men.gg+geom_smooth(method="loess")
```

Observing the trend with straight line fit.
```{r}
Age_men.gg=ggplot(Age_men_data,aes(x=Age_men_data$Age,y=Age_men_data$BPSysAve))+geom_point()
Age_men.gg+geom_smooth(method="lm")
```

Oberving the loess fit and lm fit for the age of women

loess fit
```{r}
Age_women.gg=ggplot(Age_women_data,aes(x=Age_women_data$Age,y=Age_women_data$BPSysAve))+geom_point()
Age_women.gg+geom_smooth(method="loess")
```

lm fit 
```{r}
Age_women.gg=ggplot(Age_women_data,aes(x=Age_women_data$Age,y=Age_women_data$BPSysAve))+geom_point()
Age_women.gg+geom_smooth(method="lm")
```

Plotting the residuals against the explanatory variable to check clear trends for age of men
```{r}
library(broom)
age_men.lm=lm(Age_men_data$BPSysAve~Age_men_data$Age, data=Age_men_data)
age_men.lm.df=augment(age_men.lm)
#summary(age_men.lm.df)
ggplot(age_men.lm.df,aes(x=age_men.lm.df$Age_men_data.Age+runif(nrow(age_men.lm.df),min= -0.5,max=0.5),y=age_men.lm.df$.resid+runif(nrow(age_men.lm.df),min= -0.5,max=0.5)))+geom_point(size=0.2,alpha=0.9)+geom_smooth(method="lm")+geom_abline(slope = 0, intercept = 0)
```
plotting age of men with loess fit

```{r}
age_men.leoss=loess(Age_men_data$BPSysAve~Age_men_data$Age,data=Age_men_data)
age_men.leoss.df=augment(age_men.leoss)
#summary(age_men.leoss.df)
ggplot(age_men.leoss.df,aes(x=age_men.leoss.df$Age_men_data.Age,y=age_men.leoss.df$.resid))+geom_point(size=0.2,alpha=0.9)+geom_smooth(method="loess")+geom_abline(slope=0, intercept=0)
```


plotting homoscedastity Age_men

spread_location plot

```{r}
ggplot(age_men.leoss.df,aes(x=age_men.leoss.df$.fitted,y=sqrt(abs(age_men.leoss.df$.resid))))+geom_point(size=0.2,alpha=0.9)+geom_smooth(method="loess")

```


plotting the lm fit for explanatory variable vs residuals for age of women
```{r}
age_women.lm=lm(Age_women_data$BPSysAve~Age_women_data$Age)
age_women.lm.df=augment(age_women.lm)
#summary(age_women.lm.df)
ggplot(age_women.lm.df,aes(x=age_women.lm.df$Age_women_data.Age,y=age_women.lm.df$.resid))+
  geom_point(size=0.2,alpha=0.9)+geom_smooth(method="lm")+geom_abline(slope=0,intercept=0)
```

plotting the loess fit for explanatory variable vs residuals for age of women
```{r}
age_women.leoss=loess(Age_women_data$BPSysAve~Age_women_data$Age,data=Age_women_data)
age_women.leoss.df=augment(age_women.leoss)
#summary(age_men.leoss.df)
ggplot(age_women.leoss.df,aes(x=age_women.leoss.df$Age_women_data.Age,y=age_women.leoss.df$.resid))+geom_point(size=0.2,alpha=0.9)+geom_smooth(method="loess")+geom_abline(slope=0, intercept=0)

```

Comparing the loess fit together for men and women for appropriate comparison
```{r}
men_women_data=NHANES[c("Age","BPSysAve","Gender")]
men_women_data=na.omit(men_women_data)
ggplot(men_women_data,aes(x=men_women_data$Age,y=men_women_data$BPSysAve,group=men_women_data$Gender,color=men_women_data$Gender))+geom_point(size=0.2,alpha=0.9)+geom_smooth(method="loess",method.args=list(degree=1))
```

R2 for men and women Age 
```{r}

```


Section 2 Height 

```{r}
ht_men=subset(NHANES,Gender=="male")
ht_women=subset(NHANES,Gender=="female")
ht_men_data=ht_men[c("Height","BPSysAve")]
ht_men_data=na.omit(ht_men_data)
ht_women_data=ht_women[c("Height","BPSysAve")]
ht_women_data=na.omit(ht_women_data)
```

Loess fit for ht of men with the residuals
```{r}
ht_men.loess=loess(ht_men_data$BPSysAve~ht_men_data$Height,data=ht_men_data)
ht_men.loess.df=augment(ht_men.loess)
#summary(ht_men.loess.df)
ggplot(ht_men.loess.df,aes(x=ht_men.loess.df$ht_men_data.Height,y=ht_men.loess.df$.resid))+geom_point(size=0.2,alpha=0.9)+geom_smooth(method="loess")+geom_abline(slope=0,intercept=0)+xlab("Height of men")+ylab("Residuals")
```

loess fit ht of women with residuals
```{r}
ht_women.loess=loess(ht_women_data$BPSysAve~ht_women_data$Height,data=ht_women_data)
ht_women.loess.df=augment(ht_women.loess)
#summary(ht_men.loess.df)
ggplot(ht_women.loess.df,aes(x=ht_women.loess.df$ht_women_data.Height,y=ht_women.loess.df$.resid))+geom_point(size=0.2,alpha=0.9)+geom_smooth(method="loess")+geom_abline(slope=0,intercept=0)+xlab("Height of women")+ylab("Residual")
```

Comparing lm fit residuals for women height
```{r}
ht_women.lm=lm(ht_women_data$BPSysAve~ht_women_data$Height,data=ht_women_data)
ht_women.lm.df=augment(ht_women.lm)
#summary(ht_men.loess.df)
ggplot(ht_women.lm.df,aes(x=ht_women.lm.df$ht_women_data.Height,y=ht_women.lm.df$.resid))+geom_point(size=0.2,alpha=0.9)+geom_smooth(method="lm")+geom_abline(slope=0,intercept=0)+xlab("Height of women")+ylab("Residual")
```
Comparing the quadratic fit for the ht of women 
```{r}
ht_women.lm=lm(ht_women_data$BPSysAve~ ht_women_data$Height+I(ht_women_data$Height^2),data=ht_women_data)
ht_women.lm.df=augment(ht_women.lm)
#summary(ht_men.loess.df)
ggplot(ht_women.lm.df,aes(x=ht_women.lm.df$ht_women_data.Height,y=ht_women.lm.df$.resid))+geom_point(size=0.2,alpha=0.9)+geom_smooth(method="lm")+geom_abline(slope=0,intercept=0)+xlab("Height of women")+ylab("Residual")

```


```{r}
var(ht_women.lm.df$.fitted)/var(ht_women.lm.df$ht_women_data.BPSysAve)
var(ht_women.loess.df$.fitted)/var(ht_women.loess.df$ht_women_data.BPSysAve)
var(ht_men.loess.df$.fitted)/var(ht_men.loess.df$ht_men_data.BPSysAve)
```

Comparison between the men and women height
```{r}
men_women_ht=NHANES[c("Height","BPSysAve","Gender")]
men_women_ht=na.omit(men_women_ht)
ggplot(men_women_ht,aes(x=men_women_ht$Height,y=men_women_ht$BPSysAve,group=men_women_ht$Gender,color=men_women_ht$Gender))+geom_point(size=0.2,alpha=0.9)+geom_smooth(method="loess",method.args=list(degree=1))

```

Ploting the loess plot for the Weight of men

```{r}
men_wt=subset(NHANES,Gender=="male")
women_wt=subset(NHANES,Gender=="female")
men_wt_data=men_wt[c("Weight","BPSysAve")]
men_wt_data=na.omit(men_wt_data)
women_wt_data=women_wt[c("Weight","BPSysAve")]
women_wt_data=na.omit(women_wt_data)
```

Plotting residuals against weight of men loess fit
```{r}
men_wt.loess=loess(men_wt_data$BPSysAve~men_wt_data$Weight,data=men_wt_data)
men_wt.loess.df=augment(men_wt.loess)
#summary(men_wt.loess.df)
ggplot(men_wt.loess.df,aes(x=men_wt.loess.df$men_wt_data.Weight,y=men_wt.loess.df$.resid))+geom_point(size=0.2,alpha=0.9)+geom_smooth(method="loess",method.args=list(degree=1))+geom_abline(slope=0,intercept = 0)
```

Ploting residuals against weight of women loess fit
```{r}
women_wt.loess=loess(women_wt_data$BPSysAve~women_wt_data$Weight,data=women_wt_data,family = "symmetric",degree=1)
women_wt.loess.df=augment(women_wt.loess)
ggplot(women_wt.loess.df,aes(x=women_wt.loess.df$women_wt_data.Weight,y=women_wt.loess.df$.resid))+geom_point(size=0.2,alpha=0.9)+geom_smooth(method="loess",method.args=list(degree=1))+geom_abline(slope=0,intercept = 0)
```

Plotting weights of men and women for comaprison
```{r}
men_women_wt=NHANES[c("Weight","BPSysAve","Gender")]
men_women_wt=na.omit(men_women_wt)
ggplot(men_women_wt,aes(x=men_women_wt$Weight,y=men_women_wt$BPSysAve,group=men_women_wt$Gender,color=men_women_wt$Gender))+geom_point(size=0.2,alpha=0.9)+geom_smooth(method="loess",method.args=list(degree=1))
```
```{r}
var(women_wt.loess.df$.fitted)/var(women_wt.loess.df$women_wt_data.BPSysAve)
var(men_wt.loess.df$.fitted)/var(men_wt.loess.df$men_wt_data.BPSysAve)

```

lm fit for men
```{r}
men_wt.lm=lm(men_wt_data$BPSysAve~men_wt_data$Weight,data=men_wt_data)
men_wt.lm.df=augment(men_wt.lm)
#summary(men_wt.loess.df)
ggplot(men_wt.lm.df,aes(x=men_wt.lm.df$men_wt_data.Weight,y=men_wt.lm.df$.resid))+geom_point(size=0.2,alpha=0.9)+geom_smooth(method="lm",method.args=list(degree=1))+geom_abline(slope=0,intercept = 0)

```

lm fit for weight of women
```{r}
women_wt.lm=lm(women_wt_data$BPSysAve~women_wt_data$Weight,data=women_wt_data)
women_wt.lm.df=augment(women_wt.lm)
#summary(men_wt.loess.df)
ggplot(women_wt.lm.df,aes(x=women_wt.lm.df$women_wt_data.Weight,y=women_wt.lm.df$.resid))+geom_point(size=0.2,alpha=0.9)+geom_smooth(method="lm",method.args=list(degree=1))+geom_abline(slope=0,intercept = 0)

```

R2
```{r}
var(women_wt.lm.df$.fitted)/var(women_wt.lm.df$women_wt_data.BPSysAve)
var(men_wt.lm.df$.fitted)/var(men_wt.lm.df$men_wt_data.BPSysAve)

```


Using cuberoot transform on independent variable
```{r}
men_ht_tr.loess=loess(ht_men_data$BPSysAve~I(ht_men_data$Height^(1/3)),data=ht_men_data,family="symmetric",degree=1)
men_ht_tr.loess.df=augment(men_ht_tr.loess)
#summary(men_ht_tr.loess.df)
ggplot(men_ht_tr.loess.df,aes(x=men_ht_tr.loess.df$I.ht_men_data.Height..1.3..,y=men_ht_tr.loess.df$.resid))+geom_point(size=0.2,alpha=0.9)+geom_smooth(method="loess",method.args=list(degree=1))+geom_abline(slope=0,intercept = 0)
var(men_ht_tr.loess.df$.fitted)/var(men_ht_tr.loess.df$ht_men_data.BPSysAve)
```
using the log transform for height of men vs residual
```{r}
men_ht_tr.loess=loess(ht_men_data$BPSysAve~lapply(ht_men_data$Height,function(x) log2(x)),data=ht_men_data,family="symmetric",degree=1)
men_ht_tr.loess.df=augment(men_ht_tr.loess)
#summary(men_ht_tr.loess.df)
ggplot(men_ht_tr.loess.df,aes(x=men_ht_tr.loess.df$I.ht_men_data.Height..1.3..,y=men_ht_tr.loess.df$.resid))+geom_point(size=0.2,alpha=0.9)+geom_smooth(method="loess",method.args=list(degree=1))+geom_abline(slope=0,intercept = 0)
var(men_ht_tr.loess.df$.fitted)/var(men_ht_tr.loess.df$ht_men_data.BPSysAve)
```
